---
title: Using Netlogo for Disease Modeling
author: admin
date: '2020-03-15'
slug: using-netlogo-for-disease-modeling
categories: []
tags:
  - agent-based modelling
  - netlogo
  - youtube
  - covid19
subtitle: ''
summary: ''
math: true

authors: []
lastmod: '2020-04-15T02:05:01+01:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE, echo=FALSE}
#knitr::opts_chunk$set(include = FALSE)
library(tidyverse)
library(knitr)
library(glue)
library(emo)
set.seed(1)
```


## Overview
Since the term "Simulation" or modeling was used frequently in the media recently to get an understanding of the spread of covid-19, I have created an instructive agent-based model for the simulation of disease spread and limited hospital capacities. Since this model is only "instructive" for the purpose of understanding modeling, I have also created a video to explain the model and some implications due to the simplicity of the design.

Please watch the video first, before playing around with the simulation:
English Version
```{r echo=FALSE}
blogdown::shortcode("youtube", "BSJ6r715gSk")
```

German Version
```{r echo=FALSE}
blogdown::shortcode("youtube", "2pqIvgrDcTA")
```

You can test the Simulation yourself here: http://bit.ly/NLinfection


If you are interested in seeing the current case numbers in comparison to other countries, have a look at this shiny application that I made:
https://andrecalerovaldez.shinyapps.io/CovidTimeSeries/


