---
title: Using the apastats package to write reproducible reports
author: admin
date: '2020-01-04'
slug: using-the-apastats-package-to-write-reproducible-reports
categories: []
tags:
  - stats
  - R
  - tutorial
subtitle: ''
summary: ''
math: true

authors: []
lastmod: '2020-01-04T02:05:01+01:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE, echo=FALSE}
#knitr::opts_chunk$set(include = FALSE)
library(tidyverse)
library(knitr)
library(glue)
library(emo)
set.seed(1)
```


## Overview

Writing a research report using RMarkdown allows you to start your analysis even before you have collected the data. This is especially helpful in report writing and thesis writing. When you, web-crawler, or a panel are still collecting data, you can already start writing the thesis.
However, writing results before having the data is somewhat challenging. Keep in mind that writing a discussion before having collected the data is a lot harder.

What we're going to need two things:

- a package that formats output in a way that we need it. In our case `apastats`.
- string manipulation: `stringr` and `glue`

## Installing apastats

You can install the `apastats` package using the following command:

```r
devtools::install_github('achetverikov/apastats',subdir='apastats')
```

Now, we can start. Let's assume we have already collected the following data about iris flowers `r emo::ji("wink")`.

```{r}
df <- iris %>% sample_n(5) 
df %>% kable()
```

## The analysis

### Deciding on our hypothesis

We have found in literature that the *setosa* species has a rather small petal length of only about 3-4 cm. Since we are sampling from random flowers we want to test whether our sample is also smaller than, let us say, 5cm. Thus, our null hypothesis would be $H_0: PL > 5$.

### Selecting a test

We would test this using a one-sample t-test on the subset of data.
```{r}
res <- df %>% 
  filter(Species == "setosa") %>% 
  pull(Petal.Length) %>% 
  t.test(mu = 5)
res
```

Well... This is not our final data yet, so we should not bee too happy about this result already. And this looks nothing like a report.
How can we write the test and the interpretation that it incorporates the results?

### Using apastats to describe

The `describe.ttest` function returns the correctly formatted markdown for reporting a t-test. So the following would always generate the correct output.

```{r}
tresult <- apastats::describe.ttest(res)
tresult
```

Meh. This is not really helpful either, right? We must include the text result as an inline result using the `` `r "\u0060r tresult\u0060"` ``. Then we can say:

> Setosa iris have petal leaves whose length is statistically significant different from 5 (`r tresult`).

### Preparing for both outcomes

This is nice, but we don't know whether our sample of $n=2$ actually represents the population. Therefore, we must prepare for both cases. Here the `glue` package is helpful in filling in the blanks in pre-written output.

```{r}
library(glue)


# test the p-value
alpha_level <- .05

# default is not significant
have_txt <- "do not have"
if (res$p.value < alpha_level) {
  have_txt <- "have"
} 

phrase <- glue("> Setosa iris {have_txt} petal leaves whose length is ", 
               "statistically significant different from 5 ({tresult}).")
```

This we can now output using: `` `r "\u0060r phrase\u0060"` ``.

`r phrase`

## Collecting the final data and running our analysis

Now, we can run the full script again, but with the full data set. I have omitted unnecessary output in the following code chunk and directly construct the output phrase.

```{r}
# run the test
res <- iris %>% 
  filter(Species == "setosa") %>% 
  pull(Petal.Length) %>% 
  t.test(mu = 5)

# describe results
tresult <- apastats::describe.ttest(res)

# generate phrase
have_txt <- "do not have"
if (res$p.value < alpha_level) {
  have_txt <- "have"
} 

# you can use markdown here as well
phrase <- glue("> Setosa iris {have_txt} petal leaves whose length is ",  
               "statistically significant different from 5 ({tresult}).")
```

The final outcome of our analysis now is:

`r phrase`


